The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2013-10-28 paper/ms.tex
Fixed Fig. 11 labels
New paragraph in off center emission

The model with homogeneous radiation sources allows to study the
effect of radiation sources distributed with a different symmetry than
the gas. We select photons emitted in spheres that are off-centered
with respect to the gas distribution.

Figure \ref{fig:OCspheres} shows the 
two spherical regions we define. Each has a radius of $0.5$R and can
have to possible centers positions $C$ along the $\hat{\i}$ direction:
$\pm C\hat{\i}$. We also place an observer along the $\hat{\j}$
direction, such that the emitting sphere centered on $+C\hat{\i}$ moves away
from the observer and the sphere centered on $+C\hat{\i}$ moves towards
the observer. We consider two values for $C=(0.25,0.50)R$.

@@

The main effect we found in the off-center emission is that for 
some models ($\tau = 10^6, V=100 km s^{-1}$ and $\tau = 10^7, V=200 km s^{-1}$)
the line presents three peaks (blue-shifted, centred, red-shifted), 
For models with ($\left[ \tau = 10^{5}, V=100 km s^{-1} \right]$ ,
$\left[\tau = 10^6, V=200 km s^{-1} \right]$ and $ \left[\tau = 10^{7}, 
V=300 km s^{-1}\right]$) the centred line is dominant but the presence of
the redder and bluer peak in also present which is also a new result.
Also as in the homogeneous / central model the line is broader at high 
rotational velocities.

@@

@@

The asymmetries in Groups I, II  and III cannot be fully explained by
rotation. However, some galaxies in those groups have a high flux between
the two peaks; an effect that can be induced by rotation. The same
argument applies to galaxies in the group IV in which the flux
in between of the two peaks is high ($ \geq 70 \%$ of the intensity in
the highest peak). Finally, the triple peaked lines in Galaxies in
Group V can be easily reproduced by rotation with off-centered
emission.

Another relevant sample of lines was presented by \cite{Yamada2012} 
for 91 LAEs at $z=3.1$. They find that $\sim 33 \%$ of the galaxies present 
a single peak profile. Some of the peaks have a stron symmetry, while
others do not. The symmetric single peaked progile can be explained by
rotation effects, while the assymetric cases naturally arise in
outflow/inflow models \citep{Verhamme2008, Dijkstra06}. Among the 91
LAEs, 8 reveal triple peaked profiles with a central peak and two
smaller peaks at the redder and bluer side, a feature that is present
in our off-center emission models with rotation.

The presence of single peaked profiles has been associated to
inflow/outflow dynamics \citep{XXX}. In this paper we show that gas
bulk rotation can also be considered as a probable origin for that
behaviour, provided that the observed single peak is highly
symmetric. Similarly, in the case of double peaked lines with a high
level of flux at the line center, rotation also deserveds to be considered in
the pool of possible bulk flows responsible for that feature,
specially if the two peaks have similar intensities. The case of
triple peaked lines is another clear feature of gas rotation, this
time with the additional condition of an off-centered emission.

@@

In this paper, we have estimated the effects of gas bulk rotation on
the morphology of the Lyman $\alpha$ line in star forming
galaxies. The results are based on the study of an homogeneous sphere
rotating as a solid body. We explored  a range of models by varying
the rotation speed, hydrogen optical depth, dust optical depth and
initial distribution of Ly$\alpha$ photons with respect to the gas
density. As a cross-validation, we obtaine our results from two
independently developed Monte-Carlo radiative transfer codes.

Our main result is that rotation clearly impacte the \ly line
morphology. Double peaked lines can make transitions into single
peaked lines as the rotational velocity increases, off-centered
can result into triple peaked lines.

We have quantified the results in different aspects. In Section
\ref{sec:widthpeak} we show that the line width increases up to a
factor of $2-3$ in comparison to the static case. Simultaneously, the
flux in the line center also increases with rotational velocity. This
increase induces the transition form a double peak to a single peak
line in the models where the radiation sources are homogeneusly
distributed.

>>>>

The model with homogeneous radiation sources allows to study the
effect of radiation sources distributed with a different symmetry than
the gas. We select photons emitted in spheres that are off-centred
with respect to the gas distribution.

Figure \ref{fig:OCspheres} shows the 
two spherical regions we define. Each has a radius of $0.5$R and can
have to possible centres positions $C$ along the $\hat{\i}$ direction:
$\pm C\hat{\i}$. We also place an observer along the $\hat{\j}$
direction, such that the emitting sphere centred on $+C\hat{\i}$ moves away
from the observer and the sphere centred on $+C\hat{\i}$ moves towards
the observer. We consider two values for $C=(0.25,0.50)R$.

@@

The main effect we found in the off-center emission is that for 
some models ($\tau = 10^6, V=100 km s^{-1}$ and $\tau = 10^7, V=200 km s^{-1}$)
the line presents three peaks (blue-shifted, centred, red-shifted),
to obtain clearest profiles concerning this results we run three models 
with $\tau_{\gamma}=10^{6}$, This profiles are shown in 
Fig.\ref{fig:OC_profiles_N6}, where we clearly observe the presence of 
three peaks, and also we do not find any asymmetry in the line 
produced by the off-center emission.

For models with ($[\tau = 10^{5}, V=100 km s^{-1}]$ ,
$[\tau = 10^6, V=200 km s^{-1}]$ and $ [\tau = 10^{7}, 
V=300 km s^{-1}]$) the centred line is dominant but the presence of
the redder and bluer peak in also present which is also a new result.
Also as in the homogeneous / central model the line is broader at high 
rotational velocities.

@@

\begin{figure*}
\includegraphics[scale=0.3]{NewPanel_N6_atpm5.png}
\caption{$\mathrm{Ly}\alpha$ profiles for the off-center positions.
green lines represents spheres locate at $-0,5R$ while red lines 
represents spheres at $0.5R$)
   \label{fig:OC_profiles_N6}} 
\end{figure*}

@@

The asymmetries in Groups I, II  and III cannot be fully explained by
rotation. However, some galaxies in those groups have a high flux between
the two peaks; an effect that can be induced by rotation. The same
argument applies to galaxies in the group IV in which the flux
in between of the two peaks is high ($ \geq 70 \%$ of the intensity in
the highest peak). Finally, the triple peaked lines in Galaxies in
Group V can be easily reproduced by rotation with off-centred
emission.

Another relevant sample of lines was presented by \cite{Yamada2012} 
for 91 LAEs at $z=3.1$. They find that $\sim 33 \%$ of the galaxies present 
a single peak profile. Some of the peaks have a strong symmetry, while
others do not. The symmetric single peaked profile can be explained by
rotation effects, while the asymmetric cases naturally arise in
outflow/inflow models \citep{Verhamme2008, Dijkstra06}. Among the 91
LAEs, 8 reveal triple peaked profiles with a central peak and two
smaller peaks at the redder and bluer side, a feature that is present
in our off-center emission models with rotation.

The presence of single peaked profiles has been associated to
inflow/outflow dynamics \citep{XXX}. In this paper we show that gas
bulk rotation can also be considered as a probable origin for that
behaviour, provided that the observed single peak is highly
symmetric. Similarly, in the case of double peaked lines with a high
level of flux at the line center, rotation also deserves to be considered in
the pool of possible bulk flows responsible for that feature,
specially if the two peaks have similar intensities. The case of
triple peaked lines is another clear feature of gas rotation, this
time with the additional condition of an off-centred emission.

@@

In this paper, we have estimated the effects of gas bulk rotation on
the morphology of the Lyman $\alpha$ line in star forming
galaxies. The results are based on the study of an homogeneous sphere
rotating as a solid body. We explored  a range of models by varying
the rotation speed, hydrogen optical depth, dust optical depth and
initial distribution of Ly$\alpha$ photons with respect to the gas
density. As a cross-validation, we obtain our results from two
independently developed Monte-Carlo radiative transfer codes.

Our main result is that rotation clearly impact the \ly line
morphology. Double peaked lines can make transitions into single
peaked lines as the rotational velocity increases, off-centred
can result into triple peaked lines.

We have quantified the results in different aspects. In Section
\ref{sec:widthpeak} we show that the line width increases up to a
factor of $2-3$ in comparison to the static case. Simultaneously, the
flux in the line center also increases with rotational velocity. This
increase induces the transition form a double peak to a single peak
line in the models where the radiation sources are homogeneously
distributed.
